---
title: "Apra Data Science Challenge 2025"
output: html_notebook
---

# Load Libraries
```{r}

packages <- c(
  'ggplot2','tidyverse','plotly','leaflet',
  'shiny','shinyWidgets','shinydashboard',
  'xts','forecast','TTR','treemapify',
  'DT','lubridate','RColorBrewer','scales','stopwords',
  'tidytext','stringr','wordcloud','wordcloud2','scales','dplyr','rfm',
  'SnowballC','textmineR','topicmodels','textclean','tm'
)
for (package in packages) { 
  if (!require(package, character.only = T, quietly = T)) {
    install.packages(package)
    library(package, character.only = T)
  }
}
```

# Load data
```{r}
crm <- read_csv("CRM_interacions_table.csv")
gift <- read_csv("gift_transactions_table.csv")
video <- read_csv("video_email_data_table.csv")
constituent <- read_csv("constituent_profiles_table.csv")
```

# Part 1: The Untapped Potential: Understanding Our Donor Landscape

## CRM Data Overview
```{r}
# CRM Interaction Type
g <- crm %>%
        group_by(CRM_INTERACTION_TYPE) %>%
        summarise(Total = n()) %>%
        select(CRM_INTERACTION_TYPE, Total) %>%
        ggplot(aes(x = reorder(CRM_INTERACTION_TYPE,Total) ,y = Total))  +
        geom_bar(stat = "identity",width = 0.5, fill='black')  +
        scale_y_continuous(labels = scales::comma) +
        labs(x ="CRM Interaction Type", y = "Count") + coord_flip() +
        theme(legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 12))
      
ggplotly(g)
```


### CRM Interaction Over Time
```{r}
crm <- crm %>%
        mutate(Year = lubridate::year(CRM_INTERACTION_DATE),
               Quarter = lubridate::quarter(CRM_INTERACTION_DATE),
               Month = lubridate::month(CRM_INTERACTION_DATE, label = TRUE),
               DOW = lubridate::wday(CRM_INTERACTION_DATE, label=TRUE))
```

### CRM Interaction by Year
```{r}
crm_year <- crm %>%
group_by(Year, CRM_INTERACTION_TYPE) %>%
        summarise(Total = n()) %>%
        select(Year,CRM_INTERACTION_TYPE, Total)

   g <- ggplot(crm_year, aes(as.factor(Year), Total, group=CRM_INTERACTION_TYPE, colour = CRM_INTERACTION_TYPE)) + 
      geom_line( linewidth=1) + theme_minimal() +
      labs(x = "Year", y = "Total", color="CRM Interaction Type") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

### CRM Interaction by Quarter
```{r}
crm %>%
group_by(Quarter, CRM_INTERACTION_TYPE) %>%
        summarise(Total = n()) %>%
        select(Quarter,CRM_INTERACTION_TYPE, Total) %>% 
      ggplot(aes(as.factor(Quarter), Total, group=CRM_INTERACTION_TYPE, colour = CRM_INTERACTION_TYPE)) + 
      geom_line( linewidth=1) + theme_minimal() +
      labs(x = "Quarter", y = "Total", color="CRM Interaction Type") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))

```

### CRM Interaction by Month
```{r}
crm %>%
group_by(Month, CRM_INTERACTION_TYPE) %>%
        summarise(Total = n()) %>%
        select(Month,CRM_INTERACTION_TYPE, Total) %>% 
      ggplot(aes(as.factor(Month), Total, group=CRM_INTERACTION_TYPE, colour = CRM_INTERACTION_TYPE)) + 
      geom_line( linewidth=1) + theme_minimal() +
      labs(x = "Month", y = "Total", color="CRM Interaction Type") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
```


### CRM Interaction by Day of Week
```{r}
crm %>%
group_by(DOW, CRM_INTERACTION_TYPE) %>%
        summarise(Total = n()) %>%
        select(DOW,CRM_INTERACTION_TYPE, Total) %>% 
      ggplot(aes(as.factor(DOW), Total, group=CRM_INTERACTION_TYPE, colour = CRM_INTERACTION_TYPE)) + 
      geom_line( linewidth=1) + theme_minimal() +
      labs(x = "Day of Week", y = "Total", color="CRM Interaction Type") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
```
## Gift Overview

### Gifts by CRM Interaction Type
```{r}
left_join(gift,crm,by='CONSTITUENT_ID') %>%
  group_by(CRM_INTERACTION_TYPE) %>%
  summarise(Total = sum(AMOUNT)) %>%
  select(CRM_INTERACTION_TYPE,Total) %>%
  ggplot(aes(x = reorder(CRM_INTERACTION_TYPE,Total) ,y = Total))  +
        geom_bar(stat = "identity",width = 0.5, fill='black')  +
        scale_y_continuous(labels = scales::comma) +
        labs(x ="CRM Interaction Type", y = "Donations") + coord_flip() +
        theme(legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 12))


```

### Gifts overtime
```{r}
gift <- gift %>%
        mutate(Year = lubridate::year(GIFT_DATE),
               Quarter = lubridate::quarter(GIFT_DATE),
               Month = lubridate::month(GIFT_DATE, label = TRUE),
               DOW = lubridate::wday(GIFT_DATE, label=TRUE))
```

### Gift by Year
```{r}
g <- gift %>%
group_by(Year) %>%
        summarise(Total = sum(AMOUNT)) %>%
        select(Year, Total) %>% 
        na.omit() %>%
      ggplot(aes(Year, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Year", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 90, hjust = 1))
ggplotly(g)
```

### Gift by Quarter
```{r}
g <- gift %>%
group_by(Quarter) %>%
        summarise(Total = sum(AMOUNT)) %>%
        select(Quarter, Total) %>% 
        na.omit() %>%
      ggplot(aes(Quarter, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Quarter", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

### Gift by Month
```{r}
g <- gift %>%
group_by(Month) %>%
        summarise(Total = sum(AMOUNT)) %>%
        select(Month, Total) %>% 
        na.omit() %>%
      ggplot(aes(Month, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Month", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

Gift by Day of Week
```{r}
g <- gift %>%
group_by(DOW) %>%
        summarise(Total = sum(AMOUNT)) %>%
        select(DOW, Total) %>% 
        na.omit() %>%
      ggplot(aes(DOW, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Day of Week", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

## Video Overview
### Video Views over time
```{r}
video <- video %>%
        mutate(Year = lubridate::year(SENT_DATE),
               Quarter = lubridate::quarter(SENT_DATE),
               Month = lubridate::month(SENT_DATE, label = TRUE),
               DOW = lubridate::wday(SENT_DATE, label=TRUE))
```

### Video views by year
```{r}
g <- video %>%
group_by(Year) %>%
        summarise(Total = sum(VIDEO_VIEWS)) %>%
        select(Year, Total) %>% 
        na.omit() %>%
      ggplot(aes(as.factor(Year), Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Year", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

### Video Views by Quarter
```{r}
g <- video %>%
group_by(Quarter) %>%
        summarise(Total = sum(VIDEO_VIEWS)) %>%
        select(Quarter, Total) %>% 
        na.omit() %>%
      ggplot(aes(Quarter, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Quarter", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

### Video Views by Month
```{r}
g <- video %>%
group_by(Month) %>%
        summarise(Total = sum(VIDEO_VIEWS)) %>%
        select(Month, Total) %>% 
        na.omit() %>%
      ggplot(aes(Month, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Month", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

### Video Views by Day of Week
```{r}
g <- video %>%
group_by(DOW) %>%
        summarise(Total = sum(VIDEO_VIEWS)) %>%
        select(DOW, Total) %>% 
        na.omit() %>%
      ggplot(aes(DOW, Total)) + 
      geom_bar(stat = "identity",width = 0.5, fill='black') + theme_minimal() +
      labs(x = "Day of the week", y = "Total") + 
      scale_y_continuous(labels = comma) +
      theme(legend.text = element_text(size = 10),
            legend.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text = element_text(size = 10),
            axis.text.x = element_text(angle = 0, hjust = 1))
ggplotly(g)
```

# Part 2: Donations Portfolio
## Donor RFM Analysis
```{r}

rfm_df <- gift %>%
  filter(GIFT_DATE >= '2015-01-01') %>%
  select(CONSTITUENT_ID,GIFT_DATE,AMOUNT) %>%
  na.omit()

#names(rfm_df)[names(rfm_df) == 'CONSTITUENT_ID'] <- 'customer_id'

#rfm model setup
analysis_date <- lubridate::as_date(today(), tz = "UTC")

report <- rfm_table_order(rfm_df, customer_id,GIFT_DATE,AMOUNT, analysis_date)
#segment
segment_titles <- c("First Grade", "Loyal", "Likely to be Loyal",
                    "New Ones", "Could be Promising", "Require Assistance", "Getting Less Frequent",
                    "Almost Out", "Can't Lose Them", "Don’t Show Up at All")
#numerical thresholds
 r_low <- c(4, 2, 3, 4, 3, 2, 2, 1, 1, 1)
 r_high <- c(5, 5, 5, 5, 4, 3, 3, 2, 1, 2)
 f_low <- c(4, 3, 1, 1, 1, 2, 1, 2, 4, 1)
 f_high <- c(5, 5, 3, 1, 1, 3, 2, 5, 5, 2)
 m_low <- c(4, 3, 1, 1, 1, 2, 1, 2, 4, 1)
 m_high  <- c(5, 5, 3, 1, 1, 3, 2, 5, 5, 2)

divisions<-rfm_segment(report, segment_titles, r_low, r_high, f_low, f_high, m_low, m_high)

#names(divisions)[names(divisions) == 'customer_id'] <- 'CONSTITUENT_ID'

division_count <- divisions %>% count(segment) %>% arrange(desc(n)) %>% rename(Segment = segment, Count = n)
```
```{r}

```

### RFM Metrics
```{r}
### rfm metrics - monetary
divisions %>%
group_by(segment) %>%
  summarise(Total = mean(amount)) %>%
  select(segment,Total) %>%
  ggplot(aes(x = reorder(segment,Total) ,y = Total))  +
        geom_bar(stat = "identity",width = 0.5, fill='black')  +
        scale_y_continuous(labels = scales::comma) +
        labs(x ="Segment", y = "Avg. Monetary") + coord_flip() +
        theme(legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 12))

```

```{r}
### rfm metrics - frequency
division_count %>%
group_by(Segment) %>%
  summarise(Total = mean(Count)) %>%
  select(Segment,Total) %>%
  ggplot(aes(x = reorder(Segment,Total) ,y = Total))  +
        geom_bar(stat = "identity",width = 0.5, fill='black')  +
        scale_y_continuous(labels = scales::comma) +
        labs(x ="Segment", y = "Avg. Frequency") + coord_flip() +
        theme(legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 12))

```

```{r}
### rfm metrics - recency
divisions %>%
group_by(segment) %>%
  summarise(Total = mean(recency_days)) %>%
  select(segment,Total) %>%
  ggplot(aes(x = reorder(segment,Total) ,y = Total))  +
        geom_bar(stat = "identity",width = 0.5, fill='black')  +
        scale_y_continuous(labels = scales::comma) +
        labs(x ="Segment", y = "Avg. Recency") + coord_flip() +
        theme(legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 12))
```

```{r}
## constituent ouput
### segment
### transaction, recency, monntary
```


### Donor Segmentation
```{r}
`Portfolios` <- c(unique(division_count$Segment))
ggplot(division_count, aes(area = Count, fill = `Portfolios`, label = `Segment`) ) +
  geom_treemap(stat = "identity",
  position = "identity") +
  geom_treemap_text(place = "centre",size = 12)
```

### Donor Lifetime Value
```{r}
# Calculate average donation per customer
avg_donation_per_customer <- gift %>%
  filter((GIFT_DATE >= '2015-01-01') & (GIFT_DATE < analysis_date)) %>%
  group_by(CONSTITUENT_ID) %>%
  summarize(avg_revenue = mean(AMOUNT))

# Calculate average donor lifespan (simplified, using the number of months)
avg_donor_lifespan <- gift %>%
  group_by(CONSTITUENT_ID) %>% 
  summarize(avg_lifespan = as.numeric(difftime(max(GIFT_DATE), min(GIFT_DATE), units = "days"))) %>%
  na.omit()

# Calculate CLV
clv_df <- inner_join(avg_donation_per_customer,avg_donor_lifespan,by='CONSTITUENT_ID') %>%
group_by(CONSTITUENT_ID) %>%
  mutate(CLV_calc = avg_revenue * avg_lifespan) %>%
  select(CONSTITUENT_ID,CLV_calc)


```


## Part 3: Donation Prediction

### Donor Forecasting
```{r}
# Forecast info

```
```{r}
# generating range of dates 

```
```{r}



```

```{r}

```


```{r}

```
```{r}

```

```{r}


```


```{r}
  g <- autoplot(gift_daily) +   
        scale_y_continuous(labels = scales::comma) + labs(x ="Gift Date", y = "Gift Amount",
                                                          title = "Daily Gift Chart") + 
        theme(plot.title = element_text(hjust=0.5))
      ggplotly(g)
```

```{r}
    gift_weekly <- apply.weekly(gift_xts, mean) 
      g <- autoplot(gift_weekly) +   
        scale_y_continuous(labels = scales::comma) + labs(x ="Gift Date", y = "Gift Amount",
                                                          title = "Weekly Gift Chart") +
        theme(plot.title = element_text(hjust=0.5))
      ggplotly(g)
```

```{r}
      gift_monthly <- apply.monthly(gift_xts, mean)
      g <- autoplot(gift_monthly) +   
        scale_y_continuous(labels = scales::comma) + labs(x ="Gift Date", y = "Gift Amount",
                                                          title = "Monthly Gift Chart") + 
        theme(plot.title = element_text(hjust=0.5))
      ggplotly(g)
```
```{r}
gifts_df <- gift %>%
  filter(GIFT_DATE >= '2015-01-01') %>%
  select(CONSTITUENT_ID,GIFT_DATE,AMOUNT) #%>%
  #na.omit()
```

```{r}
# --- . Data Preparation for Forecasting ---
print("Aggregating gift data to monthly totals...")
monthly_donations <- gifts_df %>%
  mutate(GIFT_DATE = ymd(GIFT_DATE)) %>%
  # Extract year and month for grouping
  mutate(year_month = floor_date(GIFT_DATE, "month")) %>%
  group_by(year_month) %>%
  summarise(total_donations = sum(AMOUNT, na.rm = TRUE), .groups = 'drop') %>%
  arrange(year_month)


if (nrow(monthly_donations) == 0) {
  stop("No monthly donation data could be aggregated. Check 'GIFT_DATE' and 'AMOUNT' columns.")
}

print(str_glue("Aggregated {nrow(monthly_donations)} monthly donation records."))
print("Sample of monthly donations:")
print(head(monthly_donations))
print(tail(monthly_donations))
```

```{r}
# --- . Create Time Series Object ---
# Determine start year and month for the time series
start_year <- lubridate::year(min(monthly_donations$year_month))
start_month <- lubridate::month(min(monthly_donations$year_month))

# Create a ts object
# Ensure no gaps in the time series if possible, or handle them
# For direct aggregation, gaps won't be explicitly in the data frame if no donations occur
# but the ts object can represent them.
donations_ts <- ts(monthly_donations$total_donations,
                    start = c(start_year, start_month),
                    frequency = 12)

print(str_glue("Created time series object starting from {start_year}-{start_month}."))
print(donations_ts)
```


```{r}
# --- . Validate Seasonality ---
png("donations_ts_plot.png", width=800, height=500)
plot(donations_ts, main = "Monthly Donations Time Series",
     xlab = "Year", ylab = "Total Donations", col = "steelblue", lwd = 2)
dev.off()
print("Time series plot saved as 'donations_ts_plot.png'.")

```

```{r}

# --- Time Series Decomposition (Additive and Multiplicative) ---
# Use additive if seasonal fluctuations are roughly constant over time
# Use multiplicative if seasonal fluctuations increase/decrease with the level of the series
tryCatch({
  decomp_add <- decompose(donations_ts, type = "additive")
  png("donations_decomp_additive.png", width=800, height=600)
  plot(decomp_add)
  dev.off()
  print("Additive decomposition plot saved as 'donations_decomp_additive.png'.")
}, error = function(e) {
  message("Could not perform additive decomposition (e.g., if series is too short): ", e$message)
})

tryCatch({
  decomp_mult <- decompose(donations_ts, type = "multiplicative")
  png("donations_decomp_multiplicative.png", width=800, height=600)
  plot(decomp_mult)
  dev.off()
  print("Multiplicative decomposition plot saved as 'donations_decomp_multiplicative.png'.")
}, error = function(e) {
  message("Could not perform multiplicative decomposition (e.g., if series is too short or has zeros/negatives): ", e$message)
})


```


```{r}
# ---Seasonal Subseries Plots ---
png("donations_seasonal_subseries_plot.png", width=800, height=500)
seasonplot(donations_ts, year.labels = TRUE, year.labels.left = TRUE,
           main = "Seasonal Subseries Plot of Monthly Donations",
           xlab = "Month", ylab = "Total Donations", col = rainbow(length(unique(year(time(donations_ts)))))
)
dev.off()
print("Seasonal subseries plot saved as 'donations_seasonal_subseries_plot.png'.")
```


```{r}
# ---STL Decomposition (Robust Decomposition)
tryCatch({
  stl_decomp <- stl(donations_ts, s.window = "periodic")
  png("donations_stl_decomp_plot.png", width=800, height=600)
  plot(stl_decomp)
  dev.off()
  print("STL decomposition plot saved as 'donations_stl_decomp_plot.png'.")

  # Print summary of STL decomposition (e.g., seasonal, trend, remainder components)
  print("\nSTL Decomposition Summary:")
  print(summary(stl_decomp))

}, error = function(e) {
  message("Could not perform STL decomposition (e.g., if series is too short): ", e$message)
})
```

```{r}
# ---Stationarity Test (Optional, for ARIMA) --
# ADF test for stationarity (null hypothesis: series has a unit root, i.e., non-stationary)
# KPSS test for stationarity (null hypothesis: series is trend-stationary)
print("\nStationarity Tests (for ARIMA suitability):")
adf_result <- adf.test(donations_ts)
print("ADF Test:")
print(adf_result)

kpss_result <- kpss.test(donations_ts)
print("KPSS Test:")
print(kpss_result)
```


```{r}

```

```{r}
# --- 4. Model Training (Automatic ARIMA) ---
print("Training an automatic ARIMA model...")
# auto.arima automatically selects the best ARIMA model
# It handles seasonality by default
arima_model <- auto.arima(donations_ts)
print("ARIMA Model Summary:")
print(summary(arima_model))
```

```{r}
png("arima_residuals_check.png", width=800, height=500)
checkresiduals(arima_model)
dev.off()
print("ARIMA residuals plot saved as 'arima_residuals_check.png'.")
```


```{r}
# ---Using ets for automatic ETS model selection
print("\nFitting ETS Model with ets:")
fit_ets <- ets(donations_ts)
print(summary(fit_ets))
```

```{r}
# --- Check residuals for white noise
png("ets_residuals_check.png", width=800, height=500)
checkresiduals(fit_ets)
dev.off()
print("ETS residuals plot saved as 'ets_residuals_check.png'.")
```


```{r}

```


```{r}
# --- 5. Forecast Future Monthly Donations ---
FORECAST_HORIZON <- 12
print(str_glue("Forecasting for the next {FORECAST_HORIZON} months..."))
forecast_arima <- forecast(arima_model, h = FORECAST_HORIZON)
forecast_ets <- forecast(fit_ets, h = FORECAST_HORIZON)

print("\nARIMA Forecast (next 12 months):")
print(forecast_arima)
print("\nETS Forecast (next 12 months):")
print(forecast_ets)
```

```{r}
# Plot forecasts
png("arima_forecast_plot.png", width=800, height=500)
plot(forecast_arima, main = "ARIMA Forecast of Monthly Donations",
     xlab = "Year", ylab = "Total Donations")
dev.off()
print("ARIMA forecast plot saved as 'arima_forecast_plot.png'.")

png("ets_forecast_plot.png", width=800, height=500)
plot(forecast_ets, main = "ETS Forecast of Monthly Donations",
     xlab = "Year", ylab = "Total Donations")
dev.off()
print("ETS forecast plot saved as 'ets_forecast_plot.png'.")
```


```{r}
print("Forecast Summary:")
print(summary(forecast_arima))

# Extracting forecast values and confidence intervals
forecast_df <- as_data_frame(forecast_arima) %>%
  rename(
    `Forecasted Donation` = `Point Forecast`,
    `Lo 80` = `Lo 80`,
    `Hi 80` = `Hi 80`,
    `Lo 95` = `Lo 95`,
    `Hi 95` = `Hi 95`
  ) %>%
  mutate(
    Month = seq(from = max(monthly_donations$year_month) + months(1),
                by = "month",
                length.out = FORECAST_HORIZON)
  ) %>%
  select(Month, `Forecasted Donation`, `Lo 80`, `Hi 80`, `Lo 95`, `Hi 95`)

print("Detailed Forecast (first 5 rows):")
print(head(forecast_df))
print("Detailed Forecast (last 5 rows):")
print(tail(forecast_df))
```


```{r}
```

```{r}
# --- Visualization ---
print("Generating forecast plot (forecast_plot.png)...")
png("forecast_plot.png", width = 800, height = 500)
plot(forecast_arima,
     main = "Monthly Donation Forecast",
     xlab = "Year",
     ylab = "Total Donations ($)",
     flty = 1, fcol = "blue")
lines(fitted(arima_model), col = "red")
legend("topleft",
       legend = c("Historical Data", "Fitted Values", "Forecasted Values", "80% CI", "95% CI"),
       col = c("black", "red", "blue", "grey", "lightgrey"),
       lty = c(1,1,1,NA,NA), pch = c(NA,NA,NA,22,22),
       pt.bg = c(NA,NA,NA,"grey","lightgrey"), bty = "n")
dev.off()
print("Forecast plot saved as 'forecast_plot.png'.")

print("--- Forecasting Model Generation Complete ---")
```

```{r}

```

```{r}
     
```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}
```

### Donor Prediction --> Next Best Donation
```{r}

```


